Distance Sampling in \proglang{R}

### 2016-01-22 paper.Rmd
More abstract updates
A few changes

  The study of plants or animals often involves the estimation of population size and/or their spatial distribution. These characteristics are important considerations in wildlife management. We introduce the \proglang{R} package \pkg{Distance} that uses distance sampling data to estimate populations size (abundance) and can be used in conjunction with another \proglang{R} package \pkg{dsm} to estimate spatial distribution. We describe example analyses using this package as well demonstrating the links with other \proglang{R} packages offering specialist solutions to specific distance sampling challenges. We demonstrate how this package provides a migration pathway from previous software, thereby allowing us to deliver cutting edge methods to the users more quickly.

@@

Conventional distance sampling assumes the observer is either stood at a point or moving along a line and will only see everything that occurs exactly at that point or on the line. It is then expected that the further an object is from the sampler (the point or line) the less likely it is that the observer will see it. By recording the distance to each of the detected objects we can fit a model to describe the probability of detection given distance from the sampler, which we refer to as the ‘detection function’.  The detection function can be used to infer how many objects were missed and thereby produce estimates of density and/or abundance. Further information on distance sampling, including its application, various extensions to deal with more complex situations and recent methodological developments can be found at http://distancesampling.org/dbib.html.

The Windows program Distance (or "DISTANCE"; for clarity henceforth "Distance for Windows") was first released in November 1998 as Distance for Windows version 3.5 and was initially based on its predecessor “TRANSECT”. Since this time it has evolved to include various design and analysis features. Distance for Windows versions 5 onwards have included \proglang{R} packages as the analysis engines providing additional more complex analysis options than those offered by the original FROTRAN engines. (LM Add part here to say what this paper describes… pathway to sustainability using R?) Further details of the history of Distance for Windows are included in @Thomas:2010cf.

>>>>

  The study of plants or animals often involves the estimation of population size and/or their spatial distribution. These characteristics are important considerations in wildlife management. We introduce the \proglang{R} package \pkg{Distance} that uses distance sampling data to estimate populations size (abundance) and is compatible with another \proglang{R} package \pkg{dsm} which can be used to model spatial distribution. We describe how users can obtain estimates of abundance and/or density using the package \pkg{Distance} as well documenting the links it provides with other \proglang{R} packages offering specialist solutions to specific distance sampling challenges. We demonstrate how this package provides a migration pathway from previous software, thereby allowing us to deliver cutting edge methods to the users more quickly.

@@

In conventional distance sampling the observer is either stood at a point or moving along a line and we only assume they see everything that occurs exactly at that point or on the line. It is then expected that the further an object is from the sampler (the point or line) the less likely it is that the observer will see it. By recording the distance to each of the detected objects we can fit a model to describe the probability of detection given distance from the sampler, which we refer to as the ‘detection function’.  The detection function can be used to infer how many objects were missed and thereby produce estimates of density and/or abundance. (LM: perhaps move this? Further information on distance sampling, including its application, various extensions to deal with more complex situations and recent methodological developments can be found at http://distancesampling.org/dbib.html.)

The Windows program Distance (or "DISTANCE"; for clarity henceforth "Distance for Windows") was first released in November 1998 as Distance for Windows version 3.5 and was initially based on its predecessor “TRANSECT”. Since this time it has evolved to include various design and analysis features. Distance for Windows versions 5 onwards have included \proglang{R} packages as the analysis engines providing additional more complex analysis options than those offered by the original FROTRAN engines. (LM to finish: The current version of Distance links to 4 R packages... Development in R progressing rapidly so allowing users to access new tools faster?)

(LM Add part here to say what this paper describes… pathway to sustainability using R?) Further details of the history of Distance for Windows are included in @Thomas:2010cf.
